Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Channel at the LHC with Multivariate Techniques

### 2015-07-14 paper/sec-mva.tex
more work on paper

%
In Fig.~\ref{fig:nnweights} we show
the distribution of the total associated weight,
Eq.~(\ref{eq:totweight}) for each of the $N_{\rm var}$ input
variables of the resolved (left plot) and boosted (right plot) categories.
%
The notation for the various kinematic variables is the same
as in Sect.~\ref{sec:input}.
%
The important information
is contained in the relative strengths of the total associated weight
for each of the input variables.
%
The results of Fig.~\ref{fig:nnweights} illustrate which variables
are more discriminative between the signal and the background.
%
For instance, in the 
resolved category, the variables that lead to
a higher discrimination power
are the $p_T$ of the two reconstructed Higgs candidates, and then on a similar
footing other variables like the $p_T$ of the individual jets
and the mass of the di-Higgs system.
%
The only variable which seems to contain to information at all
is the $\Delta \phi_{hh}$ separation between the two
Higgs candidates.

@@

For the boosted category, we see that substructure variables
are the most helpful ones, in particular the $k_t$ splitting scales
$\sqrt{d_{12}}$ and the subjetiness ratio $\tau_{12}$.
%
The ratio of energy correlations $D_2$ and $C_2$ also help,
in particular the former.
%
From the more traditional variables, the angular
separation between the two large-$R$ jets $\Delta R_{hh}$,
the invariant mass of the di-Higgs system $m_{hh}$ and the various
$p_T$ also have discrimination power.
%
Other variables provide limited information, in particular the invariant
mass of the Higgs candidates, $m_{h_1}$ and $m_{h_2}$, due to the realistic
simulation of detector resolution introduced with the smearing.
%
We have verified that signal significance in the boosted category is substantially
degraded if substructure variables are not used, emphasizing that
in the boosted category jet substructure provides the best handle
to separate signal and background.

At this point we have all the ingredients that we need to determine the optimal
value of the MVA output cut and quantify the final values of the signal
significance $S/\sqrt{B}$ and $S/B$ in our three categories.
%
These optimal values are determining by the maximisation of $S/\sqrt{B}$,
ensuring that the number of events which is left at the HL-LHC is large
enough, and also that the actual number of MC events used for the ANN training
is still large enough to avoid the biases of a small training sample.
%
The post-MVA results for $S/\sqrt{B}$ and $S/B$ as a function of the cut
in the ANN output for each of the three categories are shown in
Fig.~\ref{fig:sb_mva}.
%
As mentioned before, the optimal cuts are determined independently for each
category.
%
Note that the values of $S/\sqrt{B}$ and $S/B$ in the three cases should correspond
to the same values at the end of the cut-based analysis, as
can be verified by looking at the {\bf C2} row of Table~\ref{table:cutflow}.
%
Indeed, at the end of the cut based analysis $S/\sqrt{B}$ is approximately
1.0, 0.5 and 0.5 for the boosted, intermediate and resolved
categories, as can be seen from Fig.~\ref{fig:sb_mva}.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.48\textwidth]{plots/ssb.pdf}
\includegraphics[width=0.48\textwidth]{plots/sb.pdf}
\caption{\small
  The values of the signal significance $S/\sqrt{B}$ and of the
  signal over background ratio $S/B$ for the boosted, intermediate
  and resolved categories.
  %
  Note that the optimal cuts are determined independently for each
  category.
  %
  The no-cut values correspond to the  cut-based analysis
  results from Table~\ref{table:cutflow}.
}
\label{fig:sb_mva}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Boosted category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm mva}=0$} & \multicolumn{4}{c}{$y_{\rm mva}=0.86$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
      193   &  $3.8\,10^4$    &  0.99    &  $5.1\,10^{-3}$  & 107 & 1040 & 3.31  & 0.10\\
        \hline
  \end{tabular}
   $\,$\\
  \vspace{0.4cm}
  \noindent
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Intermediate category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm mva}=0$} & \multicolumn{4}{c}{$y_{\rm mva}=0.49$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
      660   &   $2.2\,10^6$   &   0.45   &  $3.1\,10^{-4}$ & 480  & $3.2\,10^5$ & 0.85  & $1.5\,10^{-3}$ \\
        \hline
  \end{tabular}
    $\,$\\
  \vspace{0.4cm}
  \noindent
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Resolved category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm mva}=0$} & \multicolumn{4}{c}{$y_{\rm mva}=0.17$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
    $1.6\,10^{3}$   &   $1.1\,10^{7}$   & 0.50     &  $1.5\,10^{-4}$   & 968 &
    $1.2\,10^{6}$  & 0.90 & $8.2\,10^{-4}$ \\
        \hline
  \end{tabular}
  \caption{\small The results for the number of signal and
    background events
    at the HL-LHC for both the cut-based analysis and after the
    optimal MVA cut, as well as the $S/\sqrt{B}$ and $S/B$
    values in the two cases.
    %
    From top to bottom we show the results for boosted,
    intermediate and resolved categories.
    %
    Note that the optimal value of the MVA cut
    is different in each of the three categories.
    \label{table:cutflowMVA}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%
In Fig.~\ref{fig:nnweights} we show
the distribution of the total associated weight,
Eq.~(\ref{eq:totweight}) for each of the $N_{\rm var}$ input
variables of the resolved (left plot) and boosted (right plot) categories.
%
The notation for the various kinematic variables is the same
as in Sect.~\ref{sec:input}.
%
The important information
is contained in the relative strengths of the total associated weight
for each of the input variables.
%
The results of Fig.~\ref{fig:nnweights} illustrate which variables
are more discriminative between the signal and the background.
%
For instance, in the 
resolved category, the variables that lead to
a higher discrimination power
are the $p_T$ of the two reconstructed Higgs candidates, and then on a similar
footing other variables like the $p_T$ of the individual jets
and the mass of the di-Higgs system $m_{hh}$.
%
The only variable which seems to contain to information at all
is the $\Delta \phi_{hh}$ separation between the two
Higgs candidates.

@@

For the boosted category, we see that substructure variables
are the most helpful ones, in particular the $k_t$-splitting scale
$\sqrt{d_{12}}$ and the subjettiness ratio $\tau_{12}$.
%
The ratio of ECFs $D_2^{(\beta)}$ and
$C_2^{(\beta)}$ also seems to provide
useful discrimination power.
%
Concerning other kinematic variables, the angular
separation between the two large-$R$ jets $\Delta R_{hh}$,
the invariant mass of the di-Higgs system $m_{hh}$ and the various
$p_T$ also have discrimination power.
%
Other variables provide limited information, in particular the invariant
mass of the Higgs candidates, $m_{h_1}$ and $m_{h_2}$: this
is due to the realistic
simulation of detector resolution introduced with the smearing.
%
We have verified that the post-MVA
signal significance in the boosted category is substantially
degraded if substructure variables are not used.

At this point we have all the ingredients that we need to determine the optimal
value of the MVA output cuts $y_{\rm cut}$
and quantify the final values of the signal
significance $S/\sqrt{B}$ and $S/B$ in the three categories.
%
These optimal values are determined from the maximisation of $S/\sqrt{B}$,
ensuring that the number of events $N_{\rm ev}$
which would be expected at the HL-LHC is large
enough.
%
In addition, we also require when fixing $y_{\rm cut}$
that the actual number of MC events used for the ANN training
is large enough to avoid the biases of a small training sample.

The post-MVA results for $S/\sqrt{B}$ and $S/B$ as a function of the cut
in the ANN output for each of the three categories are shown in
Fig.~\ref{fig:sb_mva}.
%
Note that the values of $S/\sqrt{B}$ and $S/B$
for $y_{\rm cut}=0$ correspond to the values at
the end of the cut-based analysis, see
the {\bf C2} row of Table~\ref{table:cutflow}.
%
Indeed, at the end of the cut based analysis $S/\sqrt{B}$ was
approximately
1.0, 0.5 and 0.5 for the boosted, intermediate and resolved
categories, as can be seen from Fig.~\ref{fig:sb_mva}.
%
The same applies to $S/B$.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.48\textwidth]{plots/ssb.pdf}
\includegraphics[width=0.48\textwidth]{plots/sb.pdf}
\caption{\small
  The values of the signal significance $S/\sqrt{B}$ and of the
  signal over background ratio $S/B$ for the boosted, intermediate
  and resolved categories as  function of the cut
  $y_{\rm cut}=0$ in the ANN output.
  %
  The $y_{\rm cut}=0$
  values correspond to the  cut-based analysis
  results from Table~\ref{table:cutflow}.
}
\label{fig:sb_mva}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Boosted category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm cut}=0$} & \multicolumn{4}{c}{$y_{\rm cut}=0.86$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
      193   &  $3.8\,10^4$    &  0.99    &  $5.1\,10^{-3}$  & 107 & 1040 & 3.31  & 0.10\\
        \hline
  \end{tabular}
   $\,$\\
  \vspace{0.4cm}
  \noindent
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Intermediate category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm cut}=0$} & \multicolumn{4}{c}{$y_{\rm cut}=0.49$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
      660   &   $2.2\,10^6$   &   0.45   &  $3.1\,10^{-4}$ & 480  & $3.2\,10^5$ & 0.85  & $1.5\,10^{-3}$ \\
        \hline
  \end{tabular}
    $\,$\\
  \vspace{0.4cm}
  \noindent
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Resolved category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm cut}=0$} & \multicolumn{4}{c}{$y_{\rm cut}=0.17$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
    $1.6\,10^{3}$   &   $1.1\,10^{7}$   & 0.50     &  $1.5\,10^{-4}$   & 968 &
    $1.2\,10^{6}$  & 0.90 & $8.2\,10^{-4}$ \\
        \hline
  \end{tabular}
  \caption{\small The results for the number of signal and
    background events
    at the HL-LHC for both the cut-based analysis and after the
    optimal MVA cut, as well as the $S/\sqrt{B}$ and $S/B$
    values in the two cases.
    %
    From top to bottom we show the results for boosted,
    intermediate and resolved categories.
    %
    The optimal value of the MVA cut $y_{\rm cut}$
    is optimised separately in the
    three categories.
    \label{table:cutflowMVA}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
